Incremental Construction of Programs in a Multitask Setting

### 2013-10-23 drafts/nips/nips2013.tex
more writeup

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

@@

\begin{figure}[h]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
  \caption{An example of a typical plate notation use. \label{fig:lda_plate}}

@@

\section{The SEC Algorithm}
\subsection{Generative Model}
Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
\subsection{Inference}
Write down EM updates. Interpret them as a form of compression solved via a weighted Neville-Manning variant.

Discuss bootstrap learning: the right curriculum allows us to build more complicated programs.
Make some analogy involving how humans learn to paint, write poetry, bake, compose music, play with blocks, or design statistical models.

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{clrscode4e}
\usepackage{tikz}
\usepackage{amsmath}

%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

@@

\begin{figure}[h]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=.7\linewidth]{./figures/lda_plate_blei.pdf}
  \caption{An example of typical plate notation use. \label{fig:lda_plate}}

@@

\section{The SEC Algorithm}
\subsection{Generative Model}
We propose a generative model over programs that captures two intuitions:
\begin{itemize}
\item{Useful programs are composed of useful subparts.}
\item{Short programs are a prior more likely.}
\end{itemize}
\begin{codebox}
\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
\li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
\li \Comment{Draw $\rho_n$, the $n^{th}$ program, by composing its subprograms, $e_n^{i}$}
\li \id{e_n^{i}} $\sim P_{e|G}(\cdot | G)$
\li \id{\rho_n} $= e_n^{\ell_n} (e_n^{\ell_n-1} ( \cdots e_n^{1}))$
\li \Comment{Draw $t_n$ by adding noise $\epsilon_n$ to the output of $\rho_n$}
\li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
\li \id{t_n} $= \rho_n + \epsilon_n$
\end{codebox}

Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
\subsection{Inference}
We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the libraray, $G$, given the tasks, $\{t_n\}$.
The joint distribution factorizes as
\begin{equation}
P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) \prod_i P(e^i_n | G)
\label{eq:joint}
\end{equation}
From equation (\ref{eq:joint}), the EM updates are
\begin{eqnarray}
q(\rho_n) &\propto& P(t_n | \rho_n) \prod_i P(e^i_n | G^{old})\\
\label{eq:qdist}
G^{new} &=& \operatorname{argmin}_G \left( -\ln P(G) -
\sum_n 
E_{q_n}
 \left[ \ln P(\rho_n | G) \right] \right)
 \label{eq:gmax}
\end{eqnarray}
Exact computation of $q(\cdot)$ in equation (\ref{eq:qdist}) would involve marginalizing over the space of all programs.
Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.

In our experiments, we used beam search. The objective function was the unnormalized posterior $P_{t_n|\rho_n}(t_n | \cdot )P_{\rho_n | G}(\cdot | G)$.
The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
Similarly, the search moves considered were those functions with the most highest prior probability under $G$.

Equation (\ref{eq:gmax}) has a natural interpretation as a form of compression.
Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.

Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
We approximate the description length of the grammar by the number of unique program fragments in the grammar, multiplied by a regularization constant.
The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
A slight generalization of the Neville-Manning algorithm then permits tractible minimization of (\ref{eq:gmax}).
Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.
